### 2020-10-26 chaloupka_sameshima_wallner_final_project.Rmd
initial commit

---
title: "Final Project"
author: "Ben Chaloupka, Tess Sameshima, and Scott Wallner"
date: "10/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```