### 2020-11-11 raw_dwell_times.Rmd
add notes to raw times
suggested additional analyses

---
title: "fractal_data_raw"
author: "sgwallner"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rio)
library(here)

```

```{r}

raw_A <- import("frac_SA.csv")

raw_B <- import("frac_SB.csv")
```

```{r}
raw_A_long <- raw_A %>% 
 pivot_longer(cols = a:f, names_to = "participant", values_to = "dwell_time")

raw_B_long <- raw_B %>% 
 pivot_longer(cols = g:p, names_to = "participant", values_to = "dwell_time")
```

```{r}
raw_A_long %>% 
  ggplot(aes(x = number, y = dwell_time, color = participant)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)

raw_B_long %>% 
  ggplot(aes(x = number, y = dwell_time, color = participant)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)

raw_A_long %>% 
  ggplot(aes(x = number, y = dwell_time, color = participant)) +
  geom_line()

raw_B_long %>% 
  ggplot(aes(x = number, y = dwell_time, color = participant)) +
  geom_line()

raw_A_long %>% 
  ggplot(aes(x = number, y = dwell_time)) +
  facet_wrap("participant") +
  geom_line() +
  geom_smooth(mehthod = "lm", se = F)

raw_B_long %>% 
  ggplot(aes(x = number, y = dwell_time)) +
  facet_wrap("participant") +
  geom_line() +
  geom_smooth(mehthod = "lm", se = F)
```

# try these analyses again with the outliers removed
# can we normalize the data
# put in pictures with tab pills, GIF would be cool
# significant difference between images
# significant difference between sequence position
# present each participants most and least viewed
# compare most and least viewed by sequence position