### 2020-11-30 presentation.Rmd
update with Tess's text and figure

---
title: "EDLD 651 Final Project Presentation"
author: "Scott Wallner, Tess Sameshima, & Ben Chaloupka"
date: "11/25/2020"
output:
  revealjs::revealjs_presentation:
    theme:  blood
    center: TRUE
    self_contained: FALSE
---

```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(revealjs)
```

# Dwelling on Complexity with Dynamic Fractals

*How much time will participants spend looking at dynamic fractals as they increase or decrease in complexity?*

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/mandy4.png" height="360" width="480">

# Ben "Turtles all the way down," Chaloupka

## Challenges faced along the way

this text will appear on the slide

## Victories and things to celebrate

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/Fractal Dwell Time Schematic.jpg" height="360" width="480">

## Challenges you are still facing

## Next R hurdle to tackle

# Tess "Best hair superlative champion," Sameshima

## Challenges faced along the way
- Technological issues 
  - Version control
  - Multi OS nomenclature
  - Error bars

 ![Alt Text](https://media.giphy.com/media/wvQIqJyNBOCjK/giphy.gif)

## Victories and things to celebrate
- Technological work arounds
  - More familiar with path settings
  - General code interpretation and debugging improvements

## Challenges you are still facing
- Fluid coding
- More error feed back comprehension

## Next R hurdle to tackle
- Using R to analyze data collected in MATLAB
- Using R for my master's project

# Scott "Dad jokes for days," Wallner

## Challenges faced along the way
- data wrangling
  - cleaning name 
  - large data sets
  - many separate data files for participants and conditions
  - needed to use both long and wide formats

## Victories and things to celebrate
- making GIFs
- super neat plots
- research question was answered
- hypothesis was correct
- using R to make this presentation

## Challenges you are still facing
- make some plots prettier

## Next R hurdle to tackle
- remove outliers and re-run analysis
- run a power analysis
- work with the "random" data
- adapt and develop the study
- run an online study
- replicate with cellular automata and 3D dynamic fractals

# Project Background and Introduction

## What are fractals?

:::: {style="display: flex;"}

:::{}
- Invented by Benoit Mandelbrot in 1980
- *"Clouds are not spheres. Mountains are not triangles."*
- Measure of "roughness" called dimension
- Self similarity and scale invariance
- Mathematically or algorithmically generated
:::

:::{}
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/cloud5.png" height="225" width="300">
:::

::::

## Where are fractals?...Everywhere!
Fractals are pervasive in natural processes.

:::: {style="display: flex;"}

::: {}
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/mountain2.png" height="200" width="300">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/purple.png" height="200" width="300">
:::

::: {}
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/veins2.png" height="200" width="300">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/tree2.png" height="200" width="300">
:::

::: {}
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/river3.png" height="200" width="300">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/Natural Fractals/mud.png" height="200" width="300">
:::

::::

## Statistical Fractal Images

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/Fractal Images Example Array.png" height="600" width="800">

## The Dwell Time Paradigm

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/Fractal Dwell Time Schematic.jpg" height="600" width="800">

## Growth, Decay and Random Presentations

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/Fractal Presentation Sequences.png" height="600" width="800">

##

<div class="col">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/4ntsf7.gif" height="300" width="400">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/4ntt1q.gif" height="300" width="400">
</div>

<div class="col">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/4ntsnw.gif" height="300" width="400">
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/4nts3v.gif" height="300" width="400">
</div>

# Substantive findings and interpretations

## 
<!-- Ben -->

## 
<!-- Tess -->
<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/ovr_dwell_time2.png" height = "300" width = "400">

## 
<!-- Scott -->

###
<small> Dwell times ***increase*** as complexity ***increases*** during the ***growth*** phase.
</small>

<small>Dwell times ***decrease*** as complexity ***decreases*** during the ***decay*** phase.
</small>

<!-- <img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/00scott.png" height="360" width="480"> -->

<img src="/Users/bchaloup/Documents/Fall 2020/EDLD651/final_project/presentation_images/00scott.png" height="480" width="640">

# Conclusions

## Thank you slide